---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
  # out.width = "100%"
)
```

# samplr: Tools To Compare Human Performance To Sampling Algorithms

<!-- badges: start -->
[![R-CMD-check](https://github.com/lucas-castillo/samplr/workflows/R-CMD-check/badge.svg)](https://github.com/lucas-castillo/samplr/actions)
<!-- badges: end -->

The goal of samplr is to provide tools to understand human performance from the perspective of sampling, both looking at how people generate samples and how people use the samples they have generated. A longer overview and other resources can be found at [sampling.warwick.ac.uk](sampling.warwick.ac.uk). Get started [here](vignettes/how-to-sample.html).

## Installation

You can install the released version of samplr from [Github](https://github.com/lucas-castillo/samplr) with:

```
devtools::install_github("lucas-castillo/samplr")
```
or alternatively use the `remotes` package 
```
remotes::install_github("lucas-castillo/samplr")
```
### Installing on MacOS
If installing on MacOS, you will need the following prior to installation:

1. Apple’s ‘Command Line Tools’: these can be (re-)installed by running `xcode-select --install` in a terminal. You may also check if those are already installed by running `pkgbuild::check_build_tools()` in R.
2. A Fortran compiler. Installers for gfortran are available [here](https://github.com/fxcoudert/gfortran-for-macOS/releases/). This installs into `/usr/local/gfortran`.

Read more about it on the [macOS Prerequisites section in the R Installation and Administration Manual](https://cran.r-project.org/doc/manuals/r-release/R-admin.html#Prerequisites).

## Example
samplr provides tools to generate samples following particular algorithms
``` {r}
library(samplr)
set.seed(1)
chain <- sampler_mh(start = 1, distr_name = "norm", distr_params = c(0,1), sigma_prop = diag(1) * .5, iterations = 2048)
print(chain[[1]][1:20])
```
As well as tools to diagnose the patterns both from samplers and participants:
```{r dpi = 300}
plot_qqplot(chain[[1]], change = TRUE)
plot_series(chain[[1]])
```


