---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
  # out.width = "100%"
)
```

# samplr: Tools To Compare Human Performance To Sampling Algorithms

<!-- badges: start -->
[![R-CMD-check](https://github.com/lucas-castillo/samplr/workflows/R-CMD-check/badge.svg)](https://github.com/lucas-castillo/samplr/actions)
[![Codecov test coverage](https://codecov.io/gh/lucas-castillo/samplr/branch/main/graph/badge.svg)](https://app.codecov.io/gh/lucas-castillo/samplr?branch=main)
 [![License: CC BY 4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)
<!-- badges: end -->

The goal of samplr is to provide tools to understand human performance from the perspective of sampling, both looking at how people generate samples and how people use the samples they have generated. A longer overview and other resources can be found at [sampling.warwick.ac.uk](sampling.warwick.ac.uk). Get started [here](vignettes/how-to-sample.html).

## Installation
You can install samplr from CRAN:
```
install.packages("samplr")
```
or install the development version from [Github](https://github.com/lucas-castillo/samplr) with:
```
devtools::install_github("lucas-castillo/samplr")
```
or alternatively using the `remotes` package 
```
remotes::install_github("lucas-castillo/samplr")
```
### Installing development version on MacOS
If you are installing the development version on MacOS, you will need the following prior to installation:

1. Apple’s ‘Command Line Tools’: these can be (re-)installed by running `xcode-select --install` in a terminal. You may also check if those are already installed by running `pkgbuild::check_build_tools()` in R.
2. A Fortran compiler. Installers for gfortran are available [here](https://github.com/fxcoudert/gfortran-for-macOS/releases/). This installs into `/usr/local/gfortran`.

Read more about it on the [macOS Prerequisites section in the R Installation and Administration Manual](https://cran.r-project.org/doc/manuals/r-release/R-admin.html#Prerequisites).

### Installing development version on Windows
If you are installing the development version on Windows, you will need to have RTools installed, which you can find [here](https://cran.r-project.org/bin/windows/Rtools/). Please make sure you install the version corresponding to your R version (i.e. for R 4.3.3, you'd need RTools 4.3).

## Example
The samplr package provides tools to generate samples following particular algorithms
``` {r, dpi = 300, fig.width=6, fig.height=3.5}
library(samplr)
set.seed(1)
chain <- sampler_mh(start = 1, distr_name = "norm", distr_params = c(0,1), sigma_prop = diag(1) * .5, iterations = 2048)
r <- plot_series(chain[[1]], change = FALSE)
```

As well as tools to diagnose the patterns both from samplers and participants:
```{r dpi = 300, fig.width=6, fig.height=5}
v <- calc_all(chain[[1]][1:200])
```
<!-- TODO: add plot_series(chain[[1]]) again -->

