---
title: "ABS Parameter Recovery"
autor: Yun-Xiao Li
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ABS Parameter Recovery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette provides the `R` scripts for parameter recovery of the Autocorrelated Bayesian sampler (ABS, Zhu et al., 2023). As ABS does not have analytical likelihood function, we employed the Approximate Bayesian Computation to conduct the parameter estimation, and the package [`abc`](https://cran.r-project.org/web/packages/abc/index.html) is required.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# ABS simulation

As a first step, we randomly generate a set of parameters and run a simulation of ABS.

```{r, message=FALSE, warning=FALSE}
require(samplr) # for the ABS
require(tidyverse)
require(abc) # for the ABC estimation
require(fdrtool) # for rhalfnorm distribution
require(ggplot2) # for ploting figures
require(parallel) # for parallel calculation
set.seed(2023)

params <- c(
  dec_bdry = 0, # the decision boundary that separates the posterior distribution into two parts
  discrim = 2, # the separation of the two distribution under the framework of Signal Detection Theory
  delta = 4, # the threshold for the difference of the numbers of samples supporting each response
  lambda = 14, # the rate of the sampling process
  nd_time = 0.3, # the lower bound of non-decision time
  s_nd_time = 0.1, # the range of non-decision time
  nChains = 10, # the number of MC3 chains
  width = 1 # the width of the proposal distribution
)

feedback <- sample(c(0,1), 200, replace = T) # the feedbacks, or say the stimuli, of each trial

ABS.sim <- rABS(params['dec_bdry'], params['discrim'], params['delta'], params['nd_time'], params['s_nd_time'],
                params['lambda'], feedback, params['nChains'], params['width'])
```

We can have an overview of the response time (RT) distributions.

```{r, fig.height=4, fig.width=7}
ggplot(data=ABS.sim, aes(x=rt))+
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(title = "Distribution of Response Time (RT)", x = "RT", y = "Frequency") +
  theme_minimal()
```

# ABC Estimation

We now consider the simulated results as empirical data and run the ABC estimation to check if we can recover the parameters. The ABC method requires summary statistics, which will be the RT quantiles, the accuracy of each stimulus, and the probability of repeats after correct and error responses in our analysis.

```{r}
summaryStats <- function(data, probs){
  
  #filter the data based on the response times
  data_clean <- filter(data, rt > 0.1 & rt < 1.5)
  
  #calculate the quantiles
  #seperate the data results based on the accuracy
  data_err <- filter(data_clean, accuracy==0)
  data_cor <- filter(data_clean, accuracy==1)
  
  #calculate the quantiles for each response
  qt_err <- quantile(data_err$rt, probs)
  qt_cor <- quantile(data_cor$rt, probs)
  
  #calculate the accuracies for each feedback
  acc_0 <- mean(data_clean$accuracy[which(data_clean$feedback == 0)])
  acc_1 <- mean(data_clean$accuracy[which(data_clean$feedback == 1)])
  
  #calculate the probability of repeats after correct and error
  prob_rept_err <- mean(data_err$rept, na.rm=T)
  prob_rept_cor <- mean(data_cor$rept, na.rm=T)
  
  return(c(qt_err, qt_cor, acc_0, acc_1, prob_rept_err, prob_rept_cor))
}

empStats <- summaryStats(ABS.sim, c(0.1, 0.3, 0.5, 0.7, 0.9))
```

## Reference table

The next step for the ABC estimation is build up the reference table, which requires generating a large number of parameter combinations, running the simulations and calculating the same summary statistics we used above for the simulation results. In this vignette we only generate 5000 parameter combinations to save time. But practically, to get a more reliable result, this number should be much higher.

```{r run-abs, cache=T}
# generate parameter combinitions and run the simulations
get_parameters <- function(i){
  c(
    dec_bdry = runif(1, min=-2, max=2), 
    discrim = runif(1, min=0, max=4),
    delta = sample(2:40, 1),
    lambda = rhalfnorm(1, theta=sqrt(pi/2)/100), 
    nd_time = runif(1, min=0.1, max=0.5),
    s_nd_time = runif(1, min=0, max=0.5),
    nChains = sample(2:10, 1),
    width = rgamma(1, 1, 1)
  )
}
params_list <- lapply(1:5000, get_parameters) # params_list is a list with the length of 10000, each of which is a set of parameters

sim_results <- lapply(params_list, function(params, feedback){
  rABS(params['dec_bdry'], params['discrim'], params['delta'], params['nd_time'], params['s_nd_time'],
                params['lambda'], feedback, params['nChains'], params['width'])}, feedback=feedback)
```

```{r}
# calculate the summary statistics for the simulations
 simStats <- sapply(sim_results, summaryStats, probs = c(0.1, 0.3, 0.5, 0.7, 0.9)) %>%
  t() %>%
  data.frame() %>%
  `colnames<-`(c('q1_err', 'q3_err', 'q5_err', 'q7_err', 'q9_err',
                 'q1_cor', 'q3_cor', 'q5_cor', 'q7_cor', 'q9_cor',
                 'acc_0', 'acc_1', 'prob_rept_err', 'prob_rept_cor'))
head(simStats)
```

If we examine the `simStats`, we will find many rows with `NA` values. It is because some certain parameter combinations produced all of the response time out of the range between 0.1 seconds and 1.5 seconds. Thus there was no valid results left after the exclusion, producing the `NA` values in the statistics. In the next step, we will remove those rows with `NA` values and the related parameter combinations.

```{r}
paramsComb <- bind_rows(params_list)
paramsComb <- paramsComb[complete.cases(simStats), ]
simStats <- simStats[complete.cases(simStats),]
```

## Cross validation

Before the parameter estimation, we should assess whether ABC is a stable method for estimating ABS by cross validation. The prediction error is defined as $$E_{pred} = \frac{\sum(\theta^{*}-\theta)^2}{n*Var(\theta)},$$ where $\theta$ is the true parameter value, $\theta^{*}$ is the predicted parameter value, and $n$ is the number of points where true and predicted values are compared. In this session, we compare two ABC method, "rejection" and "loclinear". In the "loclinear" method, we choose to apply a "log" transformation on the parameter `lambda` because the range of this parameter is much larger than the others.

```{r}
cvRej <- cv4abc(paramsComb, simStats, nval=10, tols=c(.1, .05, .025), method='rejection')
summary(cvRej)
```

```{r, warning=FALSE}
cvReg <- cv4abc(paramsComb, simStats, nval=10, tols=c(.1, .05, .025), method="loclinear", transf=c("none", "none", "none", "log", "none", "none", "none", "none"))
summary(cvReg)
```

We found that the prediction error of the parameters `nChains` and `width` are close to 1, indicating a bad cross-validation performance. In terms of the parameters `lambda` and `s_nd_time`, due to the transformation, the method of "loclinear" is better than "rejection".

## Parameter estimation

Now we estimate the posterior distribution of the parameters with the method of "loclinear".

```{r, warning=FALSE}
res <- abc(target = empStats, param = paramsComb, sumstat = simStats, tol = 0.05, method="loclinear", transf=c("none", "none", "none", "log", "none", "none", "none", "none"))
summary(res)
```

The initial values for the parameters are

| Parameters        | Values |
|-------------------|--------|
| decision boundary | 0      |
| discriminability  | 2      |
| delta             | 4      |
| lambda            | 14     |
| nd_time           | 0.3    |
| s_nd_time         | 0.1    |
| nChains           | 10     |
| width             | 1      |

Comparing the initial settings with the weighted mode of the prediction, we notice that parameter `discriminability`, `delta` and `nd_time` are recovered, but the prediction of other parameters is quite different from the initial values.

# References

-   Zhu, J.-Q., Sundh, J., Spicer, J., Chater, N., & Sanborn, A. N. (2023). The Autocorrelated Bayesian Sampler: A rational process for probability judgments, estimates, confidence intervals, choices, confidence judgments, and response times. Psychological Review. <https://doi.org/10.1037/rev0000427>
