---
title: "Custom Density Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{custom-density-functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(samplr)
```

To sample from a custom density function, create said function, then use it as the `custom_density` parameter with one of the samplers, instead of giving `distr_name`, `distr_params` (see [Supported Distributions](supported-distributions.html) and [How to Sample](how-to-sample.html)) and `weights`, if applicable.

In this example we create a 2D uniform distribution, which is not supported by Rcpp and RcppDist at the time of writing.
```{r}
# Create function
customDensity_r <- function(x){
  if (x[1] > 0 && x[1] < 3 && x[2] < 0 && x[2] > -3){
    return (1)
  } else {
    return (0)                    }
}
# Sample
Y <- sampler_mh(start = c(0,0), custom_density = customDensity_r)

# Plot results
x <- Y[[1]][,1]
y <- Y[[1]][,2]
plot(x,y, xlim = c(-5,5), ylim = c(-5,5))
```


#### On Timing

Using a custom function, rather than sampling from a supported distribution, will considerably increase the running time (as the C++ code under the hood has to constantly switch between C++ and R). Creating a density function with `Rcpp::cppFunction` doesn't change this much. 

```{r}
Rcpp::cppFunction("double customDensity_cpp(NumericVector x){
                    if (x(0) > 0 && x(0) < 3 && x(1) < 0 && x(1) > -3){
                      return 1;
                    } else {
                      return 0;
                    }
                  }")
X <- bench::mark(
  "CPP Density" = {sampler_mh(start = c(0,0), custom_density = customDensity_cpp, sigma_prop = diag(2))},
  "R Density" = {sampler_mh(start = c(0,0), custom_density = customDensity_r, sigma_prop = diag(2))},
  "Supported Density" = {sampler_mh(start = c(0,0), distr_name = "mvnorm", distr_params = list(c(0,1), diag(2)), sigma_prop = diag(2))},
  check = FALSE,
)
knitr::kable(as.data.frame(X[,c("expression", "min", "median")]))
```

Note, however, than this is still much faster than doing everything in R. 
```{r}
bespoke_mh <- function(pdf, start, iterations=1024, sigma_prop){
  # Initialize variables ---------------------------------
  acceptances <- 0
  n_dim <- length(start)
  chain <- matrix(0, nrow=iterations, ncol = n_dim)
  chain[1,] <- start
  probabilities <- c(pdf(start))

  # Run the sampler ------------------------------------------------

  for (i in 2:iterations){
    current_x <- chain[i-1,]
    # generate proposal
    proposal <- mvtnorm::rmvnorm(1, mean = current_x, sigma = sigma_prop)
    # calculate current and proposal probabilities
    prob_curr <- probabilities[i-1]
    prob_prop <- pdf(proposal)
    accept <- FALSE

    # proposal is accepted with probability prob_prop / prob_curr
    if (prob_curr != 0){
      ratio <- prob_prop / prob_curr
      if (ratio >= 1){
        accept <- TRUE
      # The beta parameter (temperature), beta <= 1, increases the value of the ratio making hotter chains more likely to accept proposals
      } else if (stats::runif(1) < ratio){
        accept <- TRUE
      }
    } else {
    # in case the current probability is 0 (in which case the ratio cannot be calculated), the step is accepted if the probability of the proposal is not 0
        if (prob_prop > 0) {
          accept <- TRUE
        }
    }
    
    if (accept){
      chain[i,] <- proposal
      acceptances <- acceptances +  1
      probabilities[i] <- prob_prop
    } else {
      chain[i,] <- current_x
      probabilities[i] <- prob_curr
    }
  
  }

  return(list(
    chain = chain,
    acceptance_ratio = acceptances/(iterations - 1)))
}

X <- bench::mark(
  "Bespoke MH" = {bespoke_mh(customDensity_r, c(0,0), iterations = 1024, sigma_prop = diag(2))}
)
knitr::kable(as.data.frame(X[,c("expression", "min", "median")]))
```
