---
title: "How to Sample From a Distribution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how-to-sample}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(samplr)
```
### Choosing a distibution
To sample from a distribution, check the abbreviation, as well as the distribution parameters, in [the Supported Distributions Vignette](supported-distributions.html). Henceforth we sample from a Normal distribution with mean 0 and standard deviation 1. So, we need:
\begin{itemize}{
  \item Distribution Name = "norm"
  \item Distribution Parameters = A vector with `mean` and `std`. In our case: `c(0,1)`. 
}

### Different Sampling Algorithms

<!-- To use a sampler from the SampleR package to explore a distribution, it is easiest to start by creating a distribution using a package such as distr or distr6.  -->
<!-- ```{r} -->
<!-- distribution <- distr::Norm(mean = 1, sd = 2) -->
<!-- ``` -->
<!-- Samplers only use the probability density function of a distribution, so the next step is to create such a function. This function will take a vector with the current position and an optional boolean with whether to return the log density instead. -->

<!-- If using distr or distr6, SampleR makes it easy to create such a function: -->
<!-- ```{r} -->
<!-- pd_function <- make_distr_pdf(distribution) -->
<!-- ``` -->
<!-- We can now ask different samplers to approximate the distribution -->
<!-- ```{r} -->
<!-- # for direct sampling, use the random function directly. -->
<!-- # these are r() for distr; $rand() for distr6; rnorm, etc, for stats, and so on  -->
<!-- direct_sampling <- distr::r(distribution)(1024) -->
<!-- # MCMC and MC3 require a starting point. We also give them the width of the proposal distribution -->
<!-- MCMC <- sampler_mcmc(pd_function, start = 0, sigma_prop = .5) -->
<!-- MC3 <- sampler_mc3(pd_function, start = 0, sigma_prop = .5) -->

<!-- # sampler_MCMC and sampler_MC3 return other information, but for now we are only -->
<!-- # interested in the regions they visited, which is the first item they return.  -->
<!-- MCMC_chain <- MCMC[[1]] -->
<!-- MC3_cold_chain <- MC3[[1]][,,1] # we are only interested in the first (cold) chain -->
<!-- ``` -->
<!-- Once we have sampled the distribution we can compare how different samplers behaved, with a series plot -->
<!-- ```{r, dpi=150, fig.align="center"} -->
<!-- a <- plot_series(direct_sampling) + ggtitle("Direct Sampling") -->
<!-- b <- plot_series(MCMC_chain) + ggtitle("MCMC") -->
<!-- c <- plot_series(MC3_cold_chain) + ggtitle("MC3") -->

<!-- gridExtra::grid.arrange(a,b,c) -->
<!-- ``` -->
<!-- A Change Series plot -->
<!-- ```{r, dpi=150, fig.align="center"} -->
<!-- a <- plot_change(direct_sampling) + ggtitle("Direct Sampling") -->
<!-- b <- plot_change(MCMC_chain) + ggtitle("MCMC") -->
<!-- c <- plot_change(MC3_cold_chain) + ggtitle("MC3") -->

<!-- gridExtra::grid.arrange(a,b,c) -->
<!-- ``` -->
<!-- And so on. A way to get a quick diagnostic look is to use the `plot_all` function -->
<!-- ```{r, fig.align="center", fig.width=10, fig.height=5} -->
<!-- plot_all(MCMC_chain, title = "MCMC") -->
<!-- ``` -->
<!-- With the other functions as a start for custom editing -->
<!-- ```{r, dpi=150, fig.align="center"} -->
<!-- plot_sigma_scaling(MCMC_chain) + ggtitle("MCMC") + theme_void() -->
<!-- ``` -->

<!-- ## A note about timing -->
<!-- When it comes to density functions, `distr6` is generally slower than `distr`. -->

<!-- ```{r} -->
<!-- chain <- stats::rnorm(2^10) -->
<!-- target <- distr::Norm() -->
<!-- target6 <- distr6::Normal$new() -->
<!-- # stats::dnorm time -->
<!-- system.time(for (i in 1:length(chain)){stats::dnorm(chain[i])}) -->
<!-- # distr time -->
<!-- system.time(for (i in 1:length(chain)){distr::d(target)(chain[i])}) -->
<!-- # distr6 time -->
<!-- system.time(for (i in 1:length(chain)){target6$pdf(chain[i])}) -->
<!-- ``` -->
<!-- However, it doesn't make much of a difference. For multivariate normals, `distr` doesn't yet have multivariate normals. We recommend using `mvtnorm` instead of `distr6` as the speed difference here is more considerable (especially in mixture distributions) -->
<!-- ```{r} -->
<!--   chain <- mvtnorm::rmvnorm(2^10, mean = c(0,0)) -->
<!--   target6 <- distr6::MultivariateNormal$new() -->
<!--   print("mvtnorm -> dmvnorm") -->
<!--   system.time(for (i in 1:length(chain[,1])){mvtnorm::dmvnorm(chain[i,])}) -->
<!--   print("distr6 -> pdf") -->
<!--   system.time(target6$pdf(data=chain)) -->
<!-- ``` -->
