---
title: "Getting Started: How to Sample From a Distribution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how-to-sample}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

```{r setup}
library(samplr)
set.seed(123)
```
### Choosing a distibution
To sample from a distribution, check the abbreviation, as well as the distribution parameters, in [the Supported Distributions Vignette](supported-distributions.html). Henceforth we sample from a Normal distribution with mean 0 and standard deviation 1. So, we need:

1. A **starting point**: Any point in the sampling space. We'll choose the mean of our distribution, `0`.,
2. **Distribution Name**: `"norm"`,
3. **Distribution Parameters**: A vector with the distribution parameters. In our case, `mean` and `std`, that is: `c(0,1)`,

```{r}
start = 0
name = "norm"
params = c(0,1)
```
### Different Sampling Algorithms
Different algorithms are available in the `samplr` package. For our example, we will use Metropolis Hastings. To use a sampler, we will always need the three previously mentioned items. For Metropolis Hastings and MC3, it is convenient to provide the variance of the proposal distribution, so we do that before sampling.
```{r}
proposal_variance <- 1 / 2
MH_samples <- sampler_mh(start, name, params, proposal_variance)
```

If we plot our results:
```{r}
hist(MH_samples[[1]], freq=FALSE, breaks = 20)
curve(dnorm(x, 0,1), add=TRUE)
```

Notice that the same approach is used for all other samplers:
```{r}
MC3_samples <- sampler_mc3(start, name, params, proposal_variance)

HMC_samples <- sampler_hmc(start, name, params)
```

### Mixture Distributions
If we want to sample from a mixture distribution, the following change:

1. **Distribution Name**: A vector of distribution names, e.g. `c("norm", "norm")`
2. **Distribution Parameters**: A list of vectors, each containing the distribution parameters.
3. **Distribution Weights (optional)**: A vector with the weights given to each distribution. The sum of weights equals 1. 

```{r}

name <- c("norm", "norm")
params <- list(c(-2, 1), c(2, 1))
weights <- c(.4, .6)
```
We'll make a custom density function to plot the curve with our results (this is not necessary for the samplers to work)
```{r}
customDensity <- function(x){
  dnorm(x, -2, 1) * 0.4 + dnorm(x, 2, 1) * 0.6
}
```
We proceed as we did before:
```{r}
# Sample (more iterations as MH struggles to change hills)
proposal_variance <- 1
MH_samples <- sampler_mh(start, name, params, proposal_variance, iterations = 2**12, weights = weights)
# plot
hist(MH_samples[[1]], freq=FALSE, breaks = 20)
curve(customDensity(x), add=TRUE)
```

See [the Multivariate Mixtures Vignette](multivariate-mixtures.html) for examples on how to mix multivariate distributions, as well as a comparison between the performance of different samplers in a 'patchy' environment. 
