---
title: "Multivariate Mixtures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multivariate-mixtures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(samplr)
library(ggplot2)
library(stats)
```

This vignette aims to show how different sampling algorithms may navigate a 'patchy' environment. Let's start by creating such an environment. 

First, we create a matrix with what will be the means of 15 different Gaussians

```{r}
set.seed(1)
# Create a matrix with the means of 15 different Gaussians
names <- rep("mvnorm", 15)
parameters <- list()
for (i in 1:15){
  parameters[[i]] <- list(runif(2) * 18 - 9, diag(2))
}
```
We'll also require a set of weights (which in this case will all be the same), which we will use in our probability density function, as well as a set of cumulative weights, for our random sampler.  
```{r} 
weights <- rep(1/15, length.out = 15)
```

The probability density function is the weighted sum of each of the density functions
```{r}
pd_func <- function(x, log = FALSE){
  densities <- vector()
  for (i in 1:length(weights)){
    densities[i] <- mvtnorm::dmvnorm(x, mean = parameters[[i]][[1]], sigma = diag(2), log = log)
  }
  return(sum(densities * weights))
}
```
<!-- The random function chooses a random distribution from the mixture to draw a sample from (with weighted probabilities) -->
<!-- ```{r} -->
<!-- rand_func <- function(x){ -->
<!--   randNums <- matrix(ncol = dims, nrow = x) -->
<!--   for (j in 1:x){ -->
<!--     randomN <- stats::runif(1) -->
<!--     finish = F -->
<!--     i = 0 -->
<!--     while (!finish){ -->
<!--       i = i + 1 -->
<!--       if (randomN <= cumulative_weights[i]){ -->
<!--         dist <- listDistr[[i]] -->
<!--         randNums[j,] <- unlist(unname(dist$rand(1))) -->
<!--         finish = T -->
<!--       } -->
<!--     } -->
<!--   } -->
<!--   return(randNums) -->
<!-- } -->
<!-- ``` -->

We can now draw a map of the density
```{r, dpi=150, fig.align="center"}
mapDensity <- function(pdf, start, size, cellsPerRow = 50){
  # start is a vector <- c(x, y)
  # size is a number n so that the map ranges from x, y to x + n, y + n
  xRange <- seq(from = start[1], to = start[1] + size, length.out = cellsPerRow)
  xxRange <- rep(xRange, cellsPerRow)

  yRange <- seq(from = start[2], to = start[2] + size, length.out = cellsPerRow)
  
  for (i in 1:cellsPerRow){
    if (i == 1){
      yyRange <- rep(yRange[i], cellsPerRow)
    } else {
      yyRange <- c(yyRange, rep(yRange[i], cellsPerRow))
    }
  }

  density <- vector()


  for (i in 1:length(yyRange)){
    density[i] <- pdf(c(xxRange[i],yyRange[i]))
    }

  df <- data.frame(x = xxRange, y = yyRange, density = density)
  return(df)
}



hills_df <- mapDensity(pd_func, c(-10,-10), 20, 150)

hill_map <- ggplot(hills_df) + 
  geom_raster(mapping = aes(x = x, y = y, fill = density)) + 
  scale_fill_viridis_c() + 
  theme_void()
print(hill_map)
```
An MCMC sampler may not be able to explore the whole of the space, as it is unable to make long jumps
```{r ,dpi=150, fig.align="center"}
iterations = 2**10
MCMC <- sampler_mcmc_cpp(
  start = c(5,5), sigma_prop = diag(2) / 8, iterations = iterations,
  distr_name = names, distr_params = parameters, discreteValues = FALSE, isMix = TRUE, weights = weights)

MCMC_df <- data.frame(x = MCMC[[1]][,1], y = MCMC[[1]][,2])
MCMC_path <- hill_map +
  geom_path(MCMC_df, mapping = aes(x,y), colour = "red", linetype = "dashed", size = .3) +
  geom_point(MCMC_df, mapping = aes(x,y), colour = "white",size =.1)
print(MCMC_path)
```
An MC3 sampler, on the other hand, runs hotter chains under the hood with which it switches stochastically, which allows it to visit far-off regions and thus explore the whole hypothesis space
```{r, dpi=150, fig.align="center"}

MC3 <-  sampler_mc3_cpp(
  start = c(5,5), nChains = 6, delta_T = 4, swap_all = FALSE,
  sigma_prop = diag(2) / 8, iterations = iterations,
  distr_name = names, distr_params = parameters,
  discreteValues = FALSE, isMix = TRUE, weights = weights
)

MC3_df <- data.frame(x = MC3[[1]][1:iterations,1], y = MC3[[1]][1:iterations,2])
MC3_path <- hill_map +
  geom_path(MC3_df, mapping = aes(x, y), colour = "red", linetype = "dashed", size = .3) +
  geom_point(MC3_df, mapping = aes(x,y), colour = "white", size =.1)
print(MC3_path)
```
